---
title: "cor_pac_guide"
author: "Tong Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cor_pac_guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction
The corr.package provides efficient and flexible tools for:

Calculating correlation coefficients using customizable methods (cor_new).
Handling missing values in datasets (handle_na).
Computing and visualizing correlation matrices for high-dimensional data (efficient_corr_matrix).
This guide demonstrates how to use these functions through detailed examples and explanations.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## getting started
```{r setup}
library(cor.package)
if (!requireNamespace("bench", quietly = TRUE)) {
    stop("The 'bench' package is required for this vignette. Please install it with install.packages('bench').")
}
library(bench)
```



## cor_new(): Flexible Correlation Calculation
The cor_new function allows for custom correlation calculations with options for different methods and handling of missing data.
```{r example-code, echo = TRUE}
# Example data
x = c(1, 2, 3, NA, 5)
y = c(2, 4, NA, 8, 10)

# Pearson correlation with mean imputation
cor_new(x, y, method = "pearson", na_handle = "mean")

# Spearman correlation with complete case handling
cor_new(x, y, method = "spearman", na_handle = "complete")

# Kendall correlation
cor_new(x, y, method = "kendall")
```


## handle_na(): Missing Value Handling
The handle_na function processes missing values in datasets based on user-defined methods.
```{r example-code, echo = TRUE}
# Example data
x = c(1, 2, 3, NA, 5)
y = c(2, 4, NA, 8, 10)

# Remove rows with missing values
handle_na(x, y, method = "complete")

# Impute missing values with the mean
handle_na(x, y, method = "mean")
```

## efficient_corr_matrix(): Correlation Matrix Calculation
The efficient_corr_matrix function computes pairwise correlations between columns of a dataset and visualizes them using a heatmap.
```{r example-code, echo = TRUE}
# Example: Compute correlation matrix for mtcars dataset
result = efficient_corr_matrix(mtcars[, 1:5], method = "pearson")

# Print the correlation matrix
result$matrix

# Visualize the correlation heatmap
result$plot
```
##Comparison Against R's cor() Function
In this section, we compare the cor_new() function with the built-in cor() function to verify both correctness and efficiency.
```{r example-code, echo = TRUE}
# Correctness Comparison
# Example data
x = c(1, 2, 3, NA, 5)
y = c(2, 4, NA, 8, 10)

# Pearson correlation with mean imputation
custom_result <- cor_new(x, y, method = "pearson", na_handle = "complete")
original_result <- cor(x, y, method = "pearson", use = "complete.obs")

# Check if the results are equal
all.equal(custom_result, original_result)
```

## Performance Comparison
```{r example-code, echo = TRUE}
# Create a large random dataset for benchmarking
set.seed(123)
x_large = rnorm(1e5)
y_large = rnorm(1e5)
# Benchmark the performance of both functions
benchmark_results = bench::mark(
  cor_new(x_large, y_large, method = "pearson", na_handle = "mean"),
  cor(x_large, y_large, method = "pearson", use = "complete.obs"),
  check = FALSE  # Avoid checking for equality during benchmarking
)

# Print benchmarking results
benchmark_results
```
## Summary
The corr.package is a powerful tool for exploring and analyzing relationships in data. With its flexible functions, you can:

Perform advanced correlation analyses with cor_new.
Preprocess datasets using handle_na.
Efficiently compute and visualize high-dimensional correlation matrices with efficient_corr_matrix.
For further information, please refer to the package documentation or help files.
