---
title: "cor_pac_guide"
author: "Tong Liu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cor_pac_guide}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: sentence
---

## Introduction

The cor.package provides efficient and flexible tools for:

Calculating correlation coefficients using customizable methods (cor_new).
Handling missing values in datasets (handle_na).
Computing and visualizing correlation matrices for high-dimensional data (efficient_corr_matrix).
Calculating correlation coefficients more efficiently using Rcpp (cor_new_rcpp_wrapper).
This guide demonstrates how to use these functions through detailed examples and explanations.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## getting started

```{r setup}
library(cor.package)
if (!requireNamespace("bench", quietly = TRUE)) {
    stop("The 'bench' package is required for this vignette. Please install it with install.packages('bench').")
}
library(bench)
```

## cor_new(): Flexible Correlation Calculation

The cor_new function allows for custom correlation calculations with options for different methods and handling of missing data.

```{r example-code_1, echo = TRUE}
# Example data
x = c(1, 2, 3, NA, 5)
y = c(2, 4, NA, 8, 10)

# Pearson correlation with mean imputation
cor_new(x, y, method = "pearson", na_handle = "mean")

# Spearman correlation with complete case handling
cor_new(x, y, method = "spearman", na_handle = "complete")

# Kendall correlation
cor_new(x, y, method = "kendall")
```

## handle_na(): Missing Value Handling

The handle_na function processes missing values in datasets based on user-defined methods.

```{r example-code_2, echo = TRUE}
# Example data
x = c(1, 2, 3, NA, 5)
y = c(2, 4, NA, 8, 10)

# Remove rows with missing values
handle_na(x, y, method = "complete")

# Impute missing values with the mean
handle_na(x, y, method = "mean")
```

## efficient_corr_matrix(): Correlation Matrix Calculation

The efficient_corr_matrix function computes pairwise correlations between columns of a dataset and visualizes them using a heatmap.

```{r example-code_3, echo = TRUE}
# Example: Compute correlation matrix for mtcars dataset
result = efficient_corr_matrix(mtcars[, 1:5], method = "pearson")

# Print the correlation matrix
result$matrix

# Visualize the correlation heatmap
result$plot
```

## Correctness test

Comparison Against R's cor() Function In this section, we compare the cor_new() function with the built-in cor() and function under C++ framework to verify both correctness and efficiency.

```{r example-code_4, echo = TRUE}
# Correctness Comparison
# Example data
x = c(1, 2, 3, NA, 5)
y = c(2, 4, NA, 8, 10)

# Pearson correlation with mean imputation
custom_result = cor_new(x, y, method = "pearson", na_handle = "complete")
original_result = cor(x, y, method = "pearson", use = "complete.obs")
rcpp_result = cor_new_rcpp_wrapper(x, y, method = "pearson", na_handle = "complete")

# Check if the results are equal
all.equal(custom_result, original_result, rcpp_result)
```

## Performance Comparison

```{r example-code_5, echo = TRUE}
# Create a large random dataset for benchmarking
rm(list = ls())
set.seed(123)
x_large = rnorm(1e5)
y_large = rnorm(1e5)
# Benchmark the performance of both functions
benchmark_results = bench::mark(
  `cor_new (R)` = cor_new(x_large, y_large, method = "pearson", na_handle = "complete"),
  `cor (base R)` = cor(x_large, y_large, method = "pearson", use = "complete.obs"),
  `cor_new_rcpp_wrapper (C++)` = cor_new_rcpp_wrapper(x_large, y_large, method = "pearson", na_handle = "complete"),
  iterations = 5,
  check = FALSE  # Avoid checking for equality during benchmarking
)

# Print benchmarking results
benchmark_results
```

## Interpretation of performance comparison

Three functions were tested for performance: cor_new, cor, and cor_new_rcpp_wrapper.
Based on the results, the in-built cor function was the fastest, while cor_new_rcpp_wrapper outperformed cor_new significantly but was still slower than the in-built cor function.
The reason for this difference is that cor_new_rcpp_wrapper, implemented in C++, still relies on loops to calculate correlation coefficients.
This design choice(choosing to use loops in C++) was made intentionally to facilitate a speed comparison between functions implemented in R and C++.

## Summary

The corr.package is a powerful tool for exploring and analyzing relationships in data.
With its flexible functions, you can:

Perform advanced correlation analyses with cor_new.
Preprocess datasets using handle_na.
Efficiently compute and visualize high-dimensional correlation matrices with efficient_corr_matrix.
For further information, please refer to the package documentation or help files.
